Feeding Dynamic Behavior into Optimization: Simulation-based Multi-objective Overtime Planning

### 2014-09-03 data/result/Article/Latex/Article.tex
.

Given F$_{R}$ and two fronts to be compared a number of quality indicators can be defined. They usually address two properties of Pareto fronts: convergence and diversity. Convergence indicators measure how close a given front F$_{A}$ is to the reference front. Diversity indicators measure the amount of solution space that is covered by a given front F$_{A}$. For the purpose of our study, we selected three indicators: Contributions (I$_{C}$), Hypervolume (I$_{HV}$), and Generational Distance $(I_{GD})$. These indicators are measured in the [0, 1] interval. I$_{C}$ is a convergence indicator that represents the proportion of solutions in a given front F$_{A}$ that also pertain to the reference front. Good Pareto fronts have high I$_{C}$, thus contributing heavily to the composition of F$_{R}$. I$_{HV}$ calculates the volume of the objective space that is covered by solutions of a given Pareto front \cite{Durillo:2009}. It mixes convergence to the reference front and diversity amongst the solutions comprising a given front. High-quality Pareto fronts present high $I_{HV}$, thus covering a large extension of the objective space. Finally, I$_{GD}$ is a convergence indicator that measures the distance between a front and the reference front \cite{vanveldhuizen:1998}. Good fronts are closer to the reference front and thus present low $I_{GD}$.

@@

On regard on I$_{C}$, NSGAII outperformed all OMS in 5 out of 6 instances. On average, each NSGAII cycle contributed with 4.7 solutions to the reference front, SH contributed with 3.3 solutions, MAR and CPM added 2.0 solutions each. Solutions generated under CPM concentrated in the parts of the solution space representing few overtime hours, a region not frequently accessed by NSGAII. All OMS added the solution representing zero overtime to the reference front. For the largest instance, SH provided more solutions than the average run of NSGAII. On regard of I$_{HV}$ and I$_{GD}$, NSGAII produced the best results for all instances. SH also provided Pareto fronts with high I$_{HV}$ and low I$_{GD}$ (as shown in the box-plots in Figure I) though not as high/low as NSGAII due to the number of vertices comprising the latter's fronts. Thus, we ratify the results presented by Ferrucci et al. \cite{Ferrucci:2013} in the sense that the CPM and margarine OMS are not competitive with optimization results. On the other hand, SH is a good approximation for NSGAII, particularly for large instances. The good performance shown by SH can be explained by the error propagation/regeneration dynamics: if overtime is concentrated towards the end of the schedule, we suppress the snowball-growth in the number of errors reaching testing activities.

@@

% Table generated by Excel2LaTeX from sheet 'tabela 4'
\begin{table}[htbp]
	\scriptsize
  \centering
  \caption{Quality indicators for each instance under our approach (NSGAII) and a model without error dynamics (NSGA$_{NE}$). Pareto fronts for NSGA$_{NE}$ cover a larger part of the solution space, leading to smaller $I_{GD}$. Solutions produced under NSGAII tend to concentrate in certain parts of the solution space}
    \begin{tabular}{llccc}
    \toprule
    \textbf{Instance} & \textbf{Config.} & \textbf{I$_{C}$} & \textbf{I$_{HV}$} & \textbf{I$_{GD}$} \\
    \midrule
    \multirow{2}[2]{*}{ACAD} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.530 $\pm$ 0.006} &         0.004 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.004  &         0.333 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.00}  \\\hline
    \multirow{2}[2]{*}{WMET} 
				  & NSGAII      &         0.020 $\pm$ 0.012  & \textbf{0.544 $\pm$ 0.004} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.003  &         0.334 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{WAMS} 
				  & NSGAII      &         0.020 $\pm$ 0.010  & \textbf{0.498 $\pm$ 0.003} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ & \textbf{0.020 $\pm$ 0.003} &         0.337 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{PSOA} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.471 $\pm$ 0.006} &         0.004 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.002  &         0.338 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{OMET} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.523 $\pm$ 0.004} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.003  &         0.337 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{PARM} 
				  & NSGAII      &         0.020 $\pm$ 0.018  & \textbf{0.462 $\pm$ 0.007} &         0.006 $\pm$ 0.001  \\
          & NSGA$_{NE}$ & \textbf{0.020 $\pm$ 0.003} &         0.338 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\
    \bottomrule
    \end{tabular}%
\end{table}%

@@

Other interesting aspect is the convex form of the Pareto fronts resulting from our formulation for the OPP. These fronts also fill the main diagonal of the hypercube representing the objective space, but instead of lines we observe curved distributions of points. These curves can be cut into regions presenting fast decrease/increase in objective values and regions on which changes in objective values are slower. For instance, consider the curves for the PARM instance in the lower part of Figure II. The cost of shrinking makespan below 350 days increases exponentially. To reduce makespan from 350 to 337 days, we have to accept an increase of about 8\% in project cost. On the other hand, reducing makespan from 360 to 350 days requires increasing the cost by only 1.2\%. On regard of overtime hours, the first reduction in makespan requires adding 38 overtime hours to the project, while the second reduction requires adding only 12 hours.

@@

One might question whether we need heuristic search to provide answers to these questions, that is, can't we find these answers relying solely on simulation? We argue that this is not possible because we are interested in relationships between different results from the simulation. If we were interested in relations between parameters and results, we could sample the selected parameters and calculate the distribution of the results under interest. But to calculate the joint distribution of two or more results we would have to sample all parameters, which is equivalent to a complete search (all combinations of possible values are assigned to parameters) or random search (a limited number of combinations are randomly picked and assigned to parameters). Complete search is unfeasible for interesting problems because the size of the solution space is $9^m$, $m$ being the number of activities in the schedule. On the other hand, results from random search are clearly outperformed by our proposed approach, as shown in RQ1. Therefore, heuristic search complements simulation by finding the points in the solution space that represent (close to) optimal combinations of simulation results (in the objective space).

Regarding the last question presented above, we observe strong concentration of overtime allocation in coding and testing activities in Pareto fronts calculated for NSGAII. This concentration is consistent with error dynamics: an increase in the number of errors produced in requirement analysis or design will snowball through the subsequent activities, increasing the cost of testing later in the schedule. Here the extensible nature of simulation provide us an interesting opportunity: one can break error propagation dynamics by using formal reviews or inspections to identify and correct errors earlier than while testing the software. The proposed simulation mechanism allows developing a scenario to represent the effects of such activities \cite{Madachy:1996} and combine them with error dynamics. The concentration of overtime in later activities is also consistent with the quality of the Pareto front built for the SH OMS. As can observed in Figure II, the Pareto front for SH has few points, but it is close to the front built for NSGAII, sometimes even showing better fitness (i.e., PARM instance). Therefore, SH can be used as a starting point for the decision-maker and can be enriched by merging its results with those calculated using our approach.

Finally, despite of having different shapes in the objective space, NSGA and NSGA$_{NE}$ present extremely high inverse correlation between the number of overtime hours and project makespan (close to -1.0 Spearman rank-order index). NSGA$_{NE}$ also presents extremely high correlation between all pairs of objectives. This leads to actionable results in the sense that algorithms might not be required to pursue optimal values for all three objectives, but only two of them: minimum makespan is observed in solutions having maximum overtime. In practical terms, discarding one objective reduces the cost of optimization, allowing for more generations to be considered and a more diverse number of solutions to be found on Pareto fronts produced by NSGAII.

>>>>

Given F$_{R}$ and two fronts to be compared a number of quality indicators can be defined. They usually address two properties of Pareto fronts: convergence and diversity. Convergence indicators measure how close a given front F$_{A}$ is to the reference front. Diversity indicators measure the amount of objective space that is covered by a given front F$_{A}$. For the purpose of our study, we selected three indicators: Contributions (I$_{C}$), Hypervolume (I$_{HV}$), and Generational Distance $(I_{GD})$. These indicators are measured in the [0, 1] interval. I$_{C}$ is a convergence indicator that represents the proportion of solutions in a given front F$_{A}$ that also pertain to the reference front. Good Pareto fronts have high I$_{C}$, thus contributing heavily to the composition of F$_{R}$. I$_{HV}$ calculates the volume of the objective space that is covered by solutions of a given Pareto front \cite{Durillo:2009}. It mixes convergence to the reference front and diversity amongst the solutions comprising a given front. High-quality Pareto fronts present high $I_{HV}$, thus covering a large extension of the objective space. Finally, I$_{GD}$ is a convergence indicator that measures the distance between a front and the reference front \cite{vanveldhuizen:1998}. Good fronts are closer to the reference front and thus present low $I_{GD}$.

@@

On regard on I$_{C}$, NSGAII outperformed all OMS in 5 out of 6 instances. On average, each NSGAII cycle contributed with 4.7 solutions to the reference front, SH contributed with 3.3 solutions, MAR and CPM added 2.0 solutions each. Solutions generated under CPM concentrated in the parts of the objective space representing few overtime hours, a region not frequently accessed by NSGAII. All OMS added the solution representing zero overtime to the reference front. For the largest instance, SH provided more solutions than the average run of NSGAII. On regard of I$_{HV}$ and I$_{GD}$, NSGAII produced the best results for all instances. SH also provided Pareto fronts with high I$_{HV}$ and low I$_{GD}$ (as shown in the box-plots in Figure I) though not as high/low as NSGAII due to the number of vertices comprising the latter's fronts. Thus, we ratify the results presented by Ferrucci et al. \cite{Ferrucci:2013} in the sense that the CPM and margarine OMS are not competitive with optimization results. On the other hand, SH is a good approximation for NSGAII, particularly for large instances. The good performance shown by SH can be explained by the error propagation/regeneration dynamics: if overtime is concentrated towards the end of the schedule, we suppress the snowball-growth in the number of errors reaching testing activities.

@@

% Table generated by Excel2LaTeX from sheet 'tabela 4'
\begin{table}[htbp]
	\scriptsize
  \centering
  \caption{Quality indicators for each instance under our approach (NSGAII) and a model without error dynamics (NSGA$_{NE}$). Pareto fronts for NSGA$_{NE}$ cover a larger part of the objective space, leading to smaller $I_{GD}$. Solutions produced under NSGAII tend to concentrate in certain parts of the objective space}
    \begin{tabular}{llccc}
    \toprule
    \textbf{Instance} & \textbf{Config.} & \textbf{I$_{C}$} & \textbf{I$_{HV}$} & \textbf{I$_{GD}$} \\
    \midrule
    \multirow{2}[2]{*}{ACAD} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.530 $\pm$ 0.006} &         0.004 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.004  &         0.333 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.00}  \\\hline
    \multirow{2}[2]{*}{WMET} 
				  & NSGAII      &         0.020 $\pm$ 0.012  & \textbf{0.544 $\pm$ 0.004} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.003  &         0.334 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{WAMS} 
				  & NSGAII      &         0.020 $\pm$ 0.010  & \textbf{0.498 $\pm$ 0.003} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ & \textbf{0.020 $\pm$ 0.003} &         0.337 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{PSOA} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.471 $\pm$ 0.006} &         0.004 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.002  &         0.338 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{OMET} 
				  & NSGAII      &         0.020 $\pm$ 0.011  & \textbf{0.523 $\pm$ 0.004} &         0.002 $\pm$ 0.001  \\
          & NSGA$_{NE}$ &         0.020 $\pm$ 0.003  &         0.337 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\\hline
    \multirow{2}[2]{*}{PARM} 
				  & NSGAII      &         0.020 $\pm$ 0.018  & \textbf{0.462 $\pm$ 0.007} &         0.006 $\pm$ 0.001  \\
          & NSGA$_{NE}$ & \textbf{0.020 $\pm$ 0.003} &         0.338 $\pm$ 0.001  & \textbf{0.001 $\pm$ 0.000} \\
    \bottomrule
    \end{tabular}%
\end{table}%

@@

Other interesting aspect is the convex form of the Pareto fronts resulting from our formulation for the OPP. These fronts also fill the main diagonal of the hypercube representing the objective space, but instead of lines we observe curved distributions of points. These curves can be cut into regions presenting fast decrease/increase in objective values and regions on which changes in objective values are slower. For instance, consider the curves for the PARM instance in the lower part of Figure II. The cost of shrinking makespan below 355 days increases faster than linearly. To reduce makespan from 355 to 345 days, we have to accept an increase of about 5.7\% in project cost. On the other hand, reducing makespan from 365 to 355 days requires increasing the cost by only 1.7\%. On regard of overtime hours, the first reduction in makespan requires adding 30 overtime hours to the project, while the second reduction requires adding only 14 hours.

@@

One might question whether we need heuristic search to answer these questions, that is, can't we find their answers relying solely on simulation? We argue that this is not possible because we are interested in relationships between different results from the simulation. If we were interested in relations between parameters and results, we could sample the selected parameters and calculate the distribution of the results. But to calculate the joint distribution of two or more results we would have to sample all parameters, which is equivalent to a complete search (all combinations of possible values are assigned to parameters) or random search (a limited number of combinations are randomly picked and assigned to parameters). Complete search is unfeasible for interesting problems due to the huge size of the objective space ($9^m$, $m$ being the number of activities). On the other hand, results from random search are outperformed by our approach, as shown in RQ1. Therefore, heuristic search complements simulation by finding the points in the objective space that represent (close to) optimal combinations of simulation results.

Regarding the last question presented above, we observe strong concentration of overtime allocation in coding and testing activities in Pareto fronts calculated for NSGAII. This concentration is consistent with error dynamics: an increase in the number of errors produced in requirement analysis or design will snowball through the subsequent activities, increasing the cost of testing later in the schedule. Here the extensible nature of simulation provides us with an interesting opportunity: one can break error propagation dynamics by using formal reviews or inspections to identify and correct errors earlier than while testing the software. The proposed simulation mechanism allows developing a scenario to represent the effects of such activities \cite{Madachy:1996} and combine them with error dynamics. The concentration of overtime in later activities is also consistent with the quality of the Pareto front built for the SH OMS. As can observed in Figure II, the Pareto front for SH has few points, but it is quite close to the front built for NSGAII. Therefore, SH can be used as a starting point for the decision-maker and can be enriched by merging its results with those calculated using our approach.

Finally, despite of having different shapes in the objective space, NSGA and NSGA$_{NE}$ present extremely high inverse correlation between the number of overtime hours and project makespan (close to -1.0 Spearman rank-order index). Indeed, NSGA$_{NE}$ presents extremely high correlation between all pairs of objectives. This leads to actionable results in the sense that algorithms might not be required to pursue optimal values for all three objectives, but only two of them: minimum makespan is observed in solutions having maximum overtime. In practical terms, discarding one objective reduces the cost of optimization, allowing for more generations to be considered and a more diverse number of solutions to be found on Pareto fronts produced by NSGAII.
